Study 2 Analysis

### 2024-10-24 study1.qmd
update with results page

---
title: "Study 1 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
---

@@

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id=ResponseId) |>
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value)

@@

# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, check values of family column
s1 |> filter(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating") |> select(family) 
# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, set values of family column to 6118
s1 <- s1 |> mutate(family=ifelse(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating",6118,family))

@@

s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator"))

>>>>

---
title: "Study 1 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: s1_hugo.md
  gfm:
    echo: true
    output-file: s1_gfm.md
---

@@

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id2=ResponseId) |>
  mutate(id=row_number()) |> relocate(id) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2)

@@

# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, check values of family column
s1 |> filter(id2=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating") |> select(family) 
# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, set values of family column to 6118
s1 <- s1 |> mutate(family=ifelse(id2=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating",6118,family))

s1 <- s1 |> 
  mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
  mutate(value=as.numeric(value))

@@

### 2024-10-24 study2.qmd
update with results page

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
---

@@

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) 
s2 <- s2 |> filter(!(id==203309)) # NA for some responses
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

s2_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=mg)) +
  geom_histogram(aes(fill=refClass), alpha=0.5, bins=20) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2") + 
  scale_y_continuous(breaks=seq(0,80,5))

s2_agg |> group_by(id,refClass,pct_goal) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=mg)) +
  geom_histogram(aes(fill=refClass), alpha=0.5, bins=20) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(pct_goal~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2") + 
  scale_y_continuous(breaks=seq(0,80,5))
```

@@

outliers <- s2_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.75)

>>>>

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
cache: true
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: s2_hugo.md
  gfm:
    echo: true
    output-file: s2_gfm.md
---

@@

#202009

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
  mutate(id=paste0("s2_",row_number())) |> relocate(id)
s2 <- s2 |> filter(!(id=="s2_78")) # NA for some responses
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
  relocate(refClass,state1,state2,calc, .after = id)

@@

s2_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=mg)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2") + 
  scale_y_continuous(breaks=seq(0,80,5))

s2_agg |> group_by(id,refClass,pct_goal) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=mg)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(pct_goal~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2") + 
  scale_y_continuous(breaks=seq(0,80,5))

@@

outliers <- s2_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>3.0)
